---
title: "Run Analysis"
author: "Wei Hao Khoong"
date: "28 May 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objectives

The purpose of this project is to demonstrate your ability to collect, work with, and clean a data set. The goal is to prepare tidy data that can be used for later analysis. You will be graded by your peers on a series of yes/no questions related to the project. You will be required to submit: 1) a tidy data set as described below, 2) a link to a Github repository with your script for performing the analysis, and 3) a code book that describes the variables, the data, and any transformations or work that you performed to clean up the data called CodeBook.md. You should also include a README.md in the repo with your scripts. This repo explains how all of the scripts work and how they are connected.

One of the most exciting areas in all of data science right now is wearable computing - see for example this article . Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The data linked to from the course website represent data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained:

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones

Here are the data for the project:

https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

```{r}
library(dplyr)
```

# Getting the Data

```{r}
#read train data
X_train <- read.table("X_train.txt")
Y_train <- read.table("Y_train.txt")
Sub_train <- read.table("subject_train.txt")
```

```{r}
#read test data
X_test <- read.table("X_test.txt")
Y_test <- read.table("Y_test.txt")
Sub_test <- read.table("subject_test.txt")

# Obtaining the Variable Descriptions & activity levels
variable_names <- read.table("features.txt")
activity_labels <- read.table("activity_labels.txt")
```

# 1. Merging the Data

```{r}
X_total <- rbind(X_train, X_test)
Y_total <- rbind(Y_train, Y_test)
Sub_total <- rbind(Sub_train, Sub_test)

```

# Extracting only the Measurements on the Mean and Standard Deviation for each Measurement

```{r}
selected_var <- variable_names[grep("mean\\(\\)|std\\(\\)",variable_names[,2]),]
X_total <- X_total[,selected_var[,1]]

```

# Using Descriptive Activity Names to Name the Activities in the Data Set

```{r}
colnames(Y_total) <- "activity"
Y_total$label <- factor(Y_total$activity, 
                        labels = as.character(activity_labels[,2]))
Y_total$activitylabel <- factor(Y_total$activity, 
                                labels = as.character(activity_labels[,2]))
activitylabel <- Y_total[,-1]
```

# Appropriately Labeling the Dataset with Descriptive Variable Names

```{r}
colnames(X_total) <- variable_names[selected_var[,1],2]
```

# Independent Tidy Dataset

```{r}
total <- cbind(X_total, Y_total)
total_mean <- total %>% group_by(label) %>% summarize_each(funs(mean))
write.table(total_mean, file = "tidydata.txt", row.names = FALSE, col.names = FALSE)
colnames(Sub_total) <- "subject"
total <- cbind(X_total, activitylabel, Sub_total)
total_mean <- total %>% group_by(activitylabel, subject) %>% summarize_each(funs(mean))
write.table(total_mean, file = "tidydata.txt", row.names = FALSE, col.names = TRUE)
```